State of the Practice for Medical Imaging Software Based on Open Source Repositories

### 2024-10-29 StateOfPractice/Papers/SOP_MI_CritReviewsBioMedEng/SOP_MI_OSR.tex
Updates up to and including Section 2

Our investigation is based on the following four research questions:
\begin{enumerate}
\item[RQ\refstepcounter{rqnum}\therqnum \label{RQ_WhatProjects}:] What MI
\textbf{open source} software projects exist? (Section~\ref{ch_results})
\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_HighestQuality}:] Based on
quantitative measurements of each project's development practices, which
projects follow best practices? (Section~\ref{ch_results})
\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareHQ2Popular}:] How
similar are the top projects identified in \rqref{RQ_HighestQuality}
to the most popular projects as viewed by the community?
(Section~\ref{Sec_VsCommunityRanking})
\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareArtifacts}:] How
do MI projects compare to general research software with respect to the
artifacts (documents, scripts and code) present in their repositories?
(Section~\ref{Sec_CompareArtifacts})
\end{enumerate}

@@

We only cover MI visualization software.  We exclude other categories of MI
software, including Segmentation, Registration, Visualization, Enhancement,
Quantification, Simulation, plus MI archiving and telemedicine systems
(Compression, Storage, and Communication).  We also exclude Statistical Analysis
and Image-based Physiological Modelling and Feature Extraction, Classification,
and Interpretation. Software that provides MI support functions is also out of
scope; therefore, we have not assessed the toolkit libraries VTK and ITK.
Finally, Picture Archiving and Communication System (PACS), which helps users to
economically store and conveniently access images, are considered out of scope.

@@

\begin{figure*}[!ht]
\includegraphics[scale=0.53]{template.pdf}
\caption{Grading template example}
\label{fg_grading_template_example}
\end{figure*}

@@

@@

The Domain Expert provided a list 12 software packages.  We found 6 packages
were on both lists: \textit{3D Slicer}, \textit{Horos}, \textit{ImageJ},
\textit{Fiji}, \textit{MRIcron} (we use its descendant \textit{MRIcroGL}) and
\textit{Mango} (we use the web version \textit{Papaya}).  The remaining six
packages were on our out-of-scope list. The Domain Expert agreed with our final
choice of 29 packages -- see Table~\ref{tab_final_list}. This contains summary
data collected in the year 2020.

The projects are sorted in descending order of lines of code.  We found the
initial release dates (Rlsd) for most projects and marked the two unknown dates
with ``?''. The date of the last update is the date of the latest update, at the
time of measurement. We found funding information (Fnd) for only eight projects.
For the Number Of Contributors (NOC) we considered anyone who made at least one
accepted commit as a contributor. The NOC is not usually the same as the number
of long-term project members, since many projects received change requests and
code from the community.  With respect to the OS, 25 packages work on all three
OSs: Windows (W), macOS (M), and Linux (L). Although the usual approach to
cross-platform compatibility was to work natively on multiple OSes, five
projects achieved platform-independence via web applications. The full
measurement data for all packages is available on [removed for blind review]
%\href{https://data.mendeley.com/datasets/k3pcdvdzj2/1} {Mendeley Data}.  
%% DOUBLE BLIND

@@

@@

In the absence of
a specific real world context, we assumed all nine qualities are equally
important. Figure~\ref{fg_overall_scores} shows the overall scores in descending
order.

\begin{figure*}[!ht]
\includegraphics[scale=0.38]{overall_scores.pdf}
\caption{Overall AHP scores with an equal weighting for all 9 software qualities}

>>>>

Our investigation is based on the following four research questions:
\begin{enumerate}
\item[RQ\refstepcounter{rqnum}\therqnum \label{RQ_WhatProjects}:] What MI
{open source} software projects exist? (Section~\ref{ch_results})
\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_HighestQuality}:] Based on
quantitative measurements of each project's development practices, which
projects follow best practices? (Section~\ref{ch_results})
\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareHQ2Popular}:] How
similar are the top projects identified in \rqref{RQ_HighestQuality}
to the most popular projects as viewed by the community?
(Section~\ref{Sec_VsCommunityRanking})
\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareArtifacts}:] How
do MI projects compare to general research software with respect to the
artifacts (documents, scripts and code) present in their repositories?
(Section~\ref{Sec_CompareArtifacts})
\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareToolsProjMngmnt}:]
How do MI projects compare to research software in general with respect to
the use of tools (Section~\ref{Sec_CompareTools}) for:
\begin{enumerate} 
	\item [\rqref{RQ_CompareToolsProjMngmnt}.a] development; and,
	\item [\rqref{RQ_CompareToolsProjMngmnt}.b] project management?
\end{enumerate}
\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareMethodologies}:]
How do MI projects compare to research software in general with respect to
principles, processes, and methodologies used?
(Section~\ref{Sec_CompareMethodologies})
\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_PainPoints}:] What are
the pain points for developers working on MI software projects?
(Section~\ref{painpoints})
\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_ComparePainPoints}:] How
do the pain points of developers from MI compare to the pain points
for research software in general? (Section~\ref{painpoints})
\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_Concerns}:] For MI
developers what specific best practices are taken to address the pain points
and software quality concerns? (Section~\ref{painpoints})
\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_Recommend}:]
What research software development practice could potentially address the
pain point concerns identified in \rqref{RQ_PainPoints}?
(Section~\ref{ch_recommendations})
\end{enumerate}

@@

We only cover MI visualization software.  We exclude other categories of MI
software, including segmentation, registration, visualization, enhancement,
quantification, simulation, plus MI archiving and telemedicine systems
(compression, storage, and communication).  We also exclude statistical analysis
and image-based physiological modelling and feature extraction, classification,
and interpretation. Software that provides MI support functions is also out of
scope; therefore, we have not assessed the toolkit libraries VTK and ITK.
Finally, Picture Archiving and Communication System (PACS), which helps users to
economically store and conveniently access images, are considered out of scope.

@@

\begin{figure*}[!ht]
\includegraphics[scale=0.67]{template.pdf}
\caption{Grading template example}
\label{fg_grading_template_example}
\end{figure*}

@@

In this section we answer \rqref{RQ_WhatProjects} and \rqref{RQ_HighestQuality}.

\subsection{In-Scope Open-Source MI Software} \label{SecWhatProjects}

@@

The Domain Expert provided a list 12 software packages.  We found 6 packages
were on both lists: \textit{3D Slicer}, \textit{Horos}, \textit{ImageJ},
\textit{Fiji}, \textit{MRIcron} (we use its descendant \textit{MRIcroGL}) and
\textit{Mango} (we use the web version \textit{Papaya}).  The remaining six
packages were on our out-of-scope list. The Domain Expert agreed with our final
choice of 29 packages.  Table~\ref{tab_final_list} summarizes the in-scope
open-source MI software that is available at the time of measurement (the year
2020), thus answering \rqref{RQ_WhatProjects}.

@@

In Table~\ref{tab_final_list} the projects are sorted in descending order of
lines of code.  We found the initial release dates (Rlsd) for most projects and
marked the two unknown dates with ``?''. The date of the last update is the date
of the latest update, at the time of measurement. We found funding information
(Fnd) for only eight projects.  For the Number Of Contributors (NOC) we
considered anyone who made at least one accepted commit as a contributor. The
NOC is not usually the same as the number of long-term project members, since
many projects received change requests and code from the community.  With
respect to the OS, 25 packages work on all three OSs: Windows (W), macOS (M),
and Linux (L). Although the usual approach to cross-platform compatibility was
to work natively on multiple OSes, five projects achieved platform-independence
via web applications. The full measurement data for all packages is available on
[removed for blind review]
%\href{https://data.mendeley.com/datasets/k3pcdvdzj2/1} {Mendeley Data}.  
%% DOUBLE BLIND

@@

\subsection{Which Projects Follow Best Practices?}

To answer \rqref{RQ_HighestQuality} we measured the software as described in
Section~\ref{SecMethodology}.  In the absence of a specific real world context,
we assumed all nine qualities are equally important.
Figure~\ref{fg_overall_scores} shows the overall AHP scores in descending order.

\begin{figure*}[!ht]
\includegraphics[scale=0.47]{overall_scores.pdf}
\caption{Overall AHP scores with an equal weighting for all 9 software qualities}
